---
title: "ESM 244 Lab 1"
author: "Keene Morrow"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# attach packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

Read in noaa_fisheries.csv data

```{r}
us_landings <-  read_csv(here::here("data", "noaa_fisheries.csv"))

```

Now, make a tidy version of the data frame.

Using:
- `str_to_lower`
- `parse_number`

```{r}
landings_tidy <- us_landings %>%
  janitor::clean_names() %>%
  mutate(state = stringr::str_to_lower(state),
         afs_name = stringr::str_to_lower(afs_name)) %>% # intentionally overwriting existing state column
  mutate(dollars_usd = parse_number(dollars_usd)) # use parse_number to remove $ from dollars_usd
```

Let's just look at salmon

Using:
- `str_remove`
- `str_detect`
- `tidyr::separate`

```{r}
salmon_landings <- landings_tidy %>%
  mutate(afs_clean = str_remove(afs_name, pattern = " aggregate")) %>% # removes string "aggregate" from column
  filter(str_detect(afs_clean, pattern = "salmon")) %>% # filters by looking for "salmon" in afs_clean
  separate(afs_clean, into = c("group", "species"), sep = ",") # splits column into two based on commas
```

Find grouped summary data


Find annual total US landings and dollar value (summing across all state) for each type of salmon using `group_by()` and `summarise()`.

```{r}
salmon_summary <- salmon_landings %>%
  group_by(year, species) %>%
  summarize(tot_landings = sum(landings_pounds),
            tot_value = sum(dollars_usd))
```

Visualization:

```{r}
salmon_landings_graph <- ggplot(data = salmon_summary,
                                aes(x = year, y = tot_landings, group = species)) + # groups species, this can also be done as color = species here or in the geom
  geom_line() +
  theme_minimal()
  

salmon_landings_graph # storing the graph as an object can make ggsave work a little better because you can call the fucntion on a specific graph, not just the last run graph
```

Exporting graph with ggsave
```{r}
ggsave(plot = salmon_landings_graph,
       here("figures", "us_salmon_km.png"),
       height = 5, width = 8)
```

Kable table:
```{r}
salmon_first_5 <- salmon_summary %>%
  head(5)
```

```{r}
kable(salmon_first_5) %>%
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)

```

